Traffic Aware Resource Allocation Schemes for Multi-Cell \acs{MIMO}-\acs{OFDM} Systems

### 2015-01-22 Report/wsrm_formulation.tex
satya_2

We now present an algorithm to solve problem \eqref{eqn-3} by using alternating optimization technique in conjunction with \acf{SCA} \cite{marks1978technical}. To do this, first by using the \ac{SINR} expression in \eqref{eq:SINR}, we equivalently reformulate problem \eqref{eqn-3} as
\iftoggle{single_column}{
\begin{IEEEeqnarray}{CCl}\label{eqn-6}
\underset{\substack{\gamma_{l,k,n},\mvec{m}{l,k,n},\\ \beta_{l,k,n},\mvec{w}{l,k,n}}}{\text{minimize}} & \quad & \|  \tilde{\mbf{v}}  \|_q \label{eqn-obj} \IEEEyessubnumber \\
\text{subject to}& \quad &\gamma_{l,k,n} \leq \frac{ | \mvec{w}{l,k,n}^\herm \mvec{H}{l,k,n} \mvec{m}{l,k,n}  |^2}{\beta_{l,k,n}} \IEEEyessubnumber \label{eqn-6.2} \\
  & \quad & \beta_{l,k,n} \geq  \enoise + \hspace{-0.75em} \sum_{(j,i) \neq (l,k)} \hspace{-0.75em} |\mvec{w}{l,k,n}^\herm \mvec{H}{b_i,k,n} \mvec{m}{j,i,n} |^2 \IEEEyessubnumber \label{eqn-6.3} \\
  & \quad & \sum_{n = 1}^N \sum_{k \in \mathcal{U}_b} \sum_{l=1}^L \text{tr} \, (\mvec{m}{l,k,n} \mvec{m}{l,k,n}^\herm) \leq P_{{\max}}, \fall b. \IEEEyessubnumber \label{eqn-6.4}
\end{IEEEeqnarray}
}{\allowdisplaybreaks
\begin{IEEEeqnarray}{CCl}\label{eqn-6}
	\underset{\substack{\gamma_{l,k,n},\mvec{m}{l,k,n},\\\beta_{l,k,n},\mvec{w}{l,k,n}}}{\text{minimize}} &\quad& \|  \tilde{\mbf{v}}  \|_q \label{eqn-obj} \IEEEyessubnumber \vspace{-0.25cm}\\
	\text{subject to} &\quad& \gamma_{l,k,n} \leq \frac{ | \mvec{w}{l,k,n}^\herm \mvec{H}{l,k,n} \mvec{m}{l,k,n}|^2}{\beta_{l,k,n}} \IEEEyessubnumber \eqspace \label{eqn-6.2} \\
	&\quad& \beta_{l,k,n} \geq \enoise + \sum_{\mathclap{(j,i) \neq (l,k)}} |\mvec{w}{l,k,n}^\herm \mvec{H}{b_i,k,n} \mvec{m}{j,i,n} |^2 \IEEEyessubnumber \eqspace \label{eqn-6.3} \\
	&\quad& \sum_{n = 1}^N \sum_{k \in \mathcal{U}_b} \sum_{l=1}^L \text{tr} \, (\mvec{m}{l,k,n} \mvec{m}{l,k,n}^\herm) \leq P_{{\max}}, \fall b. \IEEEyessubnumber \eqspace \label{eqn-6.4}
\end{IEEEeqnarray}
}
In this formulation, we relaxed the \ac{SINR} expression \eqref{eq:SINR} by the inequalities in \eqref{eqn-6.2} and \eqref{eqn-6.3}. \review{Note that \eqref{eqn-6.2} is an under estimator for \ac{SINR} \me{\gamma_{l,k,n}}, and \eqref{eqn-6.3} provides an upper bound for the total interference seen by user \me{k \in \mathcal{U}_b}, denoted by variable \me{\beta_{l,k,n}}}. \review{The constraints are tight when the system is power limited or when the total number of backlogged packets is larger than the transmission rate.}

In order to find a tractable solution for problem \eqref{eqn-6}, we note that the constraints \eqref{eqn-6.4} are convex with involved variables. Thus, we only need to deal with \eqref{eqn-6.2} and \eqref{eqn-6.3}. Towards this end, we resort to the traditional coordinate descent technique by fixing the linear receivers, and finding the optimal transmit beamformers. \review{Note that the problem in \eqref{eqn-6} is similar to the \ac{WSRM} formulation, which is known to be NP-hard even for the single antenna case \cite{np_hard}.} % Recall that the coordinate descent method assumes that the optimization variables belong to disjoint sets and the problem is convex for a variable while all other variables are fixed \cite{xu2013block}.

For a fixed receivers \me{\mvec{w}{l,k,n}}, the problem now is to find the optimal transmit beamformers \me{\mvec{m}{l,k,n}} which is still a challenging task. We note that for fixed \me{\mvec{w}{l,k,n}}, \eqref{eqn-6.3} can be written as a \ac{SOC} constraint. Thus, the difficulty is due to the non-convexity in \eqref{eqn-6.2}. \review{To arrive at a tractable formulation, we adopt \ac{SCA} to handle \eqref{eqn-6.2} by replacing the original non-convex constraint by a series of convex constraints \cite{marks1978technical}}. Let us define a function, 
\begin{equation*}
f({\mbf{u}}_{l,k,n}) \triangleq \frac{ | \mvec{w}{l,k,n}^\herm \mvec{H}{l,k,n} \mvec{m}{l,k,n}|^2}{\beta_{l,k,n}},
\end{equation*}
where \me{{\mbf{u}}_{l,k,n} \triangleq \{\mvec{w}{l,k,n}, \mvec{m}{l,k,n},\beta_{l,k,n}\}} is the vector which needs to be identified for the optimal allocation. Note that the function \me{f({\mbf{u}}_{l,k,n})} is convex for fixed \me{\mvec{w}{l,k,n}}, since it is in fact the ratio between a quadratic form of \me{\mvec{m}{l,k,n}} over an affine function of \me{\beta_{l,k,n}} \cite{boyd2004convex}. Eq. \eqref{eqn-6.2} can be viewed as a \ac{DC} constraint. Now the convex function given by \me{f({\mbf{u}}_{l,k,n})} is approximated by considering the first order approximation around the operating point \me{\tilde{\mbf{u}}_{l,k,n}}.

@@

Now, the optimal linear receivers for the fixed transmit precoders \me{\mbf{m}_{j,i,n} \, \forall i \in \mc{U}, \, \forall n \in \mc{C}} are obtained by minimizing \eqref{eqn-3} with respect to \me{\mbf{w}_{l,k,n}} as
\begin{IEEEeqnarray}{CCl}\label{eqn-9--1}
\underset{\substack{\gamma_{l,k,n},\\ \mvec{w}{l,k,n},\beta_{l,k,n}}}{\text{minimize}} &\quad & \| \tilde{\mbf{v}} \|_q \IEEEyessubnumber\label{eqn-9--1.1a} \\
\text{subject to} & \quad & \beta_{l,k,n} \geq  \enoise + \hspace{-0.75em} \sum_{(j,i) \neq (l,k)} \hspace{-0.75em} |\mvec{w}{l,k,n}^\herm \mvec{H}{b_i,k,n} \mvec{m}{j,i,n} |^2 \IEEEyessubnumber \eqspace \label{eqn-9--1.1c} \\
& \quad&\sum_{n = 1}^N \sum_{k \in \mathcal{U}_b} \sum_{l=1}^L \text{tr} \, (\mvec{m}{l,k,n} \mvec{m}{l,k,n}^\herm) \leq P_{{\max}}, \fall b, \IEEEyessubnumber \label{eqn-9--1.1d} \\
& \quad & \text{and } \eqref{eqn-8}. \IEEEyessubnumber \label{eqn-9--1.1e}
\end{IEEEeqnarray}
Solving \eqref{eqn-9--1} using the \ac{KKT} conditions, we obtain the following iterative expression for the receiver \me{\mvec{w}{l,k,n}} as
\begin{IEEEeqnarray}{rCl}
\mvec{\widetilde{R}}{l,k,n} &=& \displaystyle \sum_{\mathclap{(j,i)\neq (l,k)}} \mvec{H}{b_i,k,n} {\mbf{m}}_{j,i,n} {\mbf{m}}_{j,i,n}^\herm \mvec{H}{b_i,k,n}^\herm + N_0 \, \mathbf{I}_{N_R} \IEEEyessubnumber \\
\mvec{w}{l,k,n}^{(i)} &=& \left ( \tfrac{\tilde{\beta}_{l,k,n} {\mbf{m}}_{l,k,n}^\herm \mvec{H}{b_k,k,n}^\herm \mbf{w}_{l,k,n}^{(i-1)} }{\|\mvec{w}{l,k,n}^{(i-1)} \mvec{H}{b_k,k,n} {\mbf{m}}_{l,k,n} \|^2} \right )\mvec{\widetilde{R}}{l,k,n}^{-1}\mvec{H}{b_k,k,n} {\mbf{m}}_{l,k,n}, \IEEEyessubnumber \eqspace \label{opt-rx}
\end{IEEEeqnarray}
where \me{\mvec{w}{l,k,n}^{(i-1)}} is the receive beamformer from the earlier iteration, upon which the linear relaxation is performed for the nonconvex constraint in \eqref{eqn-9--1}. Eq. \eqref{opt-rx} is obtained by iterating over the fixed \me{\mvec{w}{l,k,n}^{(i-1)}} at each \ac{SCA} iteration until convergence or for fixed number of iterations. \review{Note that the optimal receiver has no explicit relation with the choice of the \me{q} value used in the objective function. The dependency is implied implicitly through the transmit precoders \me{\mvec{m}{l,k,n}}, which in deed depend on the \me{q} value.}

It can be seen that the optimal receiver expression in \eqref{opt-rx} is in fact a scaled version of the \ac{MMSE} receiver, which is given by
\begin{IEEEeqnarray}{rCl}
\mvec{R}{l,k,n} &=& \displaystyle \sum_{i\in \mc{U}} \sum_{j=1}^L \mvec{H}{b_i,k,n} \mvec{m}{j,i,n} \mvec{m}{j,i,n}^\herm \mvec{H}{b_i,k,n}^\herm + N_0 \, \mathbf{I}_{N_R} \IEEEyessubnumber \eqspace \\
\mvec{w}{l,k,n} &=& \mathbf{R}^{-1}_{l,k,n} \; \mvec{H}{b_k,k,n} \; \mvec{m}{l,k,n}. \IEEEyessubnumber \label{eqn-10}
\end{IEEEeqnarray}
\review{Since the scaling present in the optimal receiver \eqref{opt-rx} has no impact on the received \acp{SINR}, the \ac{MMSE} receiver in \eqref{eqn-10} can also be used without compromising the performance or the convergence behavior}.

>>>>

We now present an algorithm to solve problem \eqref{eqn-3} by using alternating optimization technique in conjunction with \acf{SCA} \cite{marks1978technical}. To do this, first by using the \ac{SINR} expression in \eqref{eq:SINR}, we equivalently reformulate problem \eqref{eqn-3} as
\iftoggle{single_column}{
\begin{IEEEeqnarray}{CCl}\label{eqn-6}
\underset{\substack{\gamma_{l,k,n},\mvec{m}{l,k,n},\\ \beta_{l,k,n},\mvec{w}{l,k,n}}}{\text{minimize}} & \quad & \|  \tilde{\mbf{v}}  \|_q \label{eqn-obj} \IEEEyessubnumber \\
\text{subject to}& \quad &\gamma_{l,k,n} \leq \frac{ | \mvec{w}{l,k,n}^\herm \mvec{H}{l,k,n} \mvec{m}{l,k,n}  |^2}{\beta_{l,k,n}} \IEEEyessubnumber \label{eqn-6.2} \\
  & \quad & \beta_{l,k,n} \geq  \enoise + \hspace{-0.75em} \sum_{(j,i) \neq (l,k)} \hspace{-0.75em} |\mvec{w}{l,k,n}^\herm \mvec{H}{b_i,k,n} \mvec{m}{j,i,n} |^2 \IEEEyessubnumber \label{eqn-6.3} \\
  & \quad & \sum_{n = 1}^N \sum_{k \in \mathcal{U}_b} \sum_{l=1}^L \text{tr} \, (\mvec{m}{l,k,n} \mvec{m}{l,k,n}^\herm) \leq P_{{\max}}, \fall b. \IEEEyessubnumber \label{eqn-6.4}
\end{IEEEeqnarray}
}{\allowdisplaybreaks
\begin{IEEEeqnarray}{CCl}\label{eqn-6}
	\underset{\substack{\gamma_{l,k,n},\mvec{m}{l,k,n},\\\beta_{l,k,n},\mvec{w}{l,k,n}}}{\text{minimize}} &\quad& \|  \tilde{\mbf{v}}  \|_q \label{eqn-obj} \IEEEyessubnumber \vspace{-0.25cm}\\
	\text{subject to} &\quad& \gamma_{l,k,n} \leq \frac{ | \mvec{w}{l,k,n}^\herm \mvec{H}{l,k,n} \mvec{m}{l,k,n}|^2}{\beta_{l,k,n}} \IEEEyessubnumber \eqspace \label{eqn-6.2} \\
	&\quad& \beta_{l,k,n} \geq \enoise + \sum_{\mathclap{(j,i) \neq (l,k)}} |\mvec{w}{l,k,n}^\herm \mvec{H}{b_i,k,n} \mvec{m}{j,i,n} |^2 \IEEEyessubnumber \eqspace \label{eqn-6.3} \\
	&\quad& \sum_{n = 1}^N \sum_{k \in \mathcal{U}_b} \sum_{l=1}^L \text{tr} \, (\mvec{m}{l,k,n} \mvec{m}{l,k,n}^\herm) \leq P_{{\max}}, \fall b. \IEEEyessubnumber \eqspace \label{eqn-6.4}
\end{IEEEeqnarray}
}
In this formulation, we relaxed the \ac{SINR} expression \eqref{eq:SINR} by the inequalities in \eqref{eqn-6.2} and \eqref{eqn-6.3}. \review{Note that \eqref{eqn-6.2} is an under estimator for \ac{SINR} \me{\gamma_{l,k,n}}, and \eqref{eqn-6.3} provides an upper bound for the total interference seen by user \me{k \in \mathcal{U}_b}, denoted by variable \me{\beta_{l,k,n}}}. \review{The constraints are tight when the per \ac{BS} objective is non zero at the optimal point, as discussed in Appendix \ref{a-3}}. \review{Note that the problem in \eqref{eqn-6} is similar to the \ac{WSRM} formulation, which is known to be NP-hard even for the single antenna case \cite{np_hard}.}

In order to find a tractable solution for problem \eqref{eqn-6}, we note that the constraints \eqref{eqn-6.4} are convex with involved variables. Thus, we only need to deal with \eqref{eqn-6.2} and \eqref{eqn-6.3}. Towards this end, we resort to the traditional coordinate descent technique by fixing the linear receivers, and finding the optimal transmit beamformers. For a fixed receivers \me{\mvec{w}{l,k,n}}, the problem now is to find the optimal transmit beamformers \me{\mvec{m}{l,k,n}} which is still a challenging task. We note that for fixed \me{\mvec{w}{l,k,n}}, \eqref{eqn-6.3} can be written as a \ac{SOC} constraint. Thus, the difficulty is due to the non-convexity in \eqref{eqn-6.2}. \review{To arrive at a tractable formulation, we adopt \ac{SCA} to handle \eqref{eqn-6.2} by replacing the original non-convex constraint by a series of convex constraints \cite{marks1978technical}}. Let us define a function, 
\begin{equation*}
f({\mbf{u}}_{l,k,n}) \triangleq \frac{ | \mvec{w}{l,k,n}^\herm \mvec{H}{l,k,n} \mvec{m}{l,k,n}|^2}{\beta_{l,k,n}},
\end{equation*}
where \me{{\mbf{u}}_{l,k,n} \triangleq \{\mvec{w}{l,k,n}, \mvec{m}{l,k,n},\beta_{l,k,n}\}} is the vector which needs to be identified for the optimal allocation. Note that the function \me{f({\mbf{u}}_{l,k,n})} is convex for fixed \me{\mvec{w}{l,k,n}}, since it is in fact the ratio between a quadratic form of \me{\mvec{m}{l,k,n}} over an affine function of \me{\beta_{l,k,n}} \cite{boyd2004convex}. \review{Eq. \eqref{eqn-6.2} can be viewed as a \ac{DC} constraint, in which the convex function given by \me{f({\mbf{u}}_{l,k,n})} is upper bounded by first order approximation around an operating point \me{\tilde{\mbf{u}}_{l,k,n}}}.

@@

Now, the optimal linear receivers for the fixed transmit precoders \me{\mbf{m}_{j,i,n} \, \forall i \in \mc{U}, \, \forall n \in \mc{C}} are obtained by minimizing \eqref{eqn-3} with respect to \me{\mbf{w}_{l,k,n}} as
\begin{IEEEeqnarray}{CCl}\label{eqn-9--1}
\underset{\substack{\gamma_{l,k,n},\\ \mvec{w}{l,k,n},\beta_{l,k,n}}}{\text{minimize}} &\quad & \| \tilde{\mbf{v}} \|_q \IEEEyessubnumber\label{eqn-9--1.1a} \\
\text{subject to} & \quad & \beta_{l,k,n} \geq  \enoise + \hspace{-0.75em} \sum_{(j,i) \neq (l,k)} \hspace{-0.75em} |\mvec{w}{l,k,n}^\herm \mvec{H}{b_i,k,n} \mvec{m}{j,i,n} |^2 \IEEEyessubnumber \eqspace \label{eqn-9--1.1c} \\
& \quad & \text{and } \eqref{eqn-8}. \IEEEyessubnumber \label{eqn-9--1.1e}
\end{IEEEeqnarray}
Solving \eqref{eqn-9--1} using the \ac{KKT} conditions, we obtain the following iterative expression for the receiver \me{\mbf{w}^{\ast}_{l,k,n}} as
\begin{IEEEeqnarray}{rCl}
\mvec{{A}}{l,k,n} &=& \displaystyle \sum_{\mathclap{(j,i)\neq (l,k)}} \mvec{H}{b_i,k,n} {\mbf{m}}_{j,i,n} {\mbf{m}}_{j,i,n}^\herm \mvec{H}{b_i,k,n}^\herm + N_0 \, \mathbf{I}_{N_R} \IEEEyessubnumber \\
\mvec{w}{l,k,n}^{(i)} &=& \left ( \tfrac{\tilde{\beta}_{l,k,n} {\mbf{m}}_{l,k,n}^\herm \mvec{H}{b_k,k,n}^\herm \mbf{w}_{l,k,n}^{(i-1)} }{\|\mvec{w}{l,k,n}^{(i-1)} \mvec{H}{b_k,k,n} {\mbf{m}}_{l,k,n} \|^2} \right )\mvec{{A}}{l,k,n}^{-1}\mvec{H}{b_k,k,n} {\mbf{m}}_{l,k,n}, \IEEEyessubnumber \eqspace \label{opt-rx}
\end{IEEEeqnarray}
where \me{\mvec{w}{l,k,n}^{(i-1)}} is the receive beamformer from the previous iteration, upon which the linear relaxation is performed for the nonconvex constraint in \eqref{eqn-9--1}. \review{The optimal receiver \me{\mbf{w}^{\ast}_{l,k,n}} is obtained by either iterating \eqref{opt-rx} until convergence or for fixed number of iterations.} \review{Note that the receiver has no explicit relation with the choice of \me{\ell_q} norm used in the objective function. The dependency is implicitly implied by the transmit precoders \me{\mvec{m}{l,k,n}}, which in deed depend on the \me{q} value.}

It can be seen that the optimal receiver in \eqref{opt-rx} is in fact a scaled version of the \ac{MMSE} receiver, which is given by
\begin{IEEEeqnarray}{rCl}
\mvec{R}{l,k,n} &=& \displaystyle \sum_{i\in \mc{U}} \sum_{j=1}^L \mvec{H}{b_i,k,n} \mvec{m}{j,i,n} \mvec{m}{j,i,n}^\herm \mvec{H}{b_i,k,n}^\herm + N_0 \, \mathbf{I}_{N_R} \IEEEyessubnumber \eqspace \\
\mvec{w}{l,k,n} &=& \mathbf{R}^{-1}_{l,k,n} \; \mvec{H}{b_k,k,n} \; \mvec{m}{l,k,n}. \IEEEyessubnumber \label{eqn-10}
\end{IEEEeqnarray}
\review{Since the scaling present in the optimal receiver \eqref{opt-rx} has no impact on the received \acp{SINR}, the \ac{MMSE} receiver in \eqref{eqn-10} can also be used without compromising the performance or the convergence behavior}.
